\documentclass[a4paper,12pt]{article}
\usepackage[T1]{fontenc}
\usepackage [isolatin]{inputenc}
\usepackage{graphicx}
\usepackage{listings}

\begin{document}

\title{\emph{Hyde Documentation}}
\author{Nicolas de S\'er\'eville and {\'A}ngel Miguel S{\'a}nchez Ben\'{\i}tez}

\maketitle 
\pagebreak
\tableofcontents
\pagebreak


\section{Introduction}
The Hyde project is developed within the NPTool framework. For the 
moment only the tracker of charged particles is currently under study.
Coupling the tracker with a gamma-ray calorimeter such as AGATA or
PARIS will be considered in the near future. 

NPTool is a modular package allowing to perform Geant4 simulations and to 
analyse the results of the simulations. It is strongly encouraged to read 
the general NPTool documentation that you can find in this directory.


\section{NPSimulation}
\subsection{Specificity of Hyde}
The Hyde tracker detector, even if it is made of several detectors of
different shapes (square, trapezoid, annular, ...), is considered as {\it one} 
detector from the NPSimulation point of view. The Hyde tracker detector 
is described in the HydeTracker class defined in the 
HydeTracker.\{hh,cc\} files. Since the Hyde tracker detector is 
registered in the DetectorConstructor.cc file it is available for NPSimulation.

In order to manage the different detector shapes (square, trapezoid, annular, 
...) of the Hyde tracker, the HydeTracker class holds a vector of 
HydeTrackerModule objects from which are, and should be, deriving all 
the different shapes (HydeTrackerSquare, HydeTrackerAnnular,
HydeTrackerTrapezoid and HydeTrackerDummyShape classes).


\subsection{Running the simulation}
To run NPSimulation the following command line should be executed: 

\begin{verbatim}
   Simulation -E xxx.reaction -D yyy.detector
\end{verbatim}

where xxx.reaction is an input file describing the event generator and
yyy.detector is an input file describing the detector geometry. All these
input files are based on keywords and can be found in the 
\$NPTool/Inputs subdirectories.

\subsubsection{Event Generators}
All the different kind of event generator files as well as their 
respective keywords are described in the general NPTool documentation. 

\subsubsection{Detector Configurations}
The keywords associated to the detector geometry file are different for 
each detector. In case of the Hyde tracker detector an example with 
all the detector shapes available at the moment is given in the following:

\begin{verbatim}
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   HydeTracker
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   HYDAnnular
           Z=      -156.5
           RMIN=   16
           RMAX=   52
           FIRSTSTAGE= 1
           SECONDSTAGE= 0
           THIRDSTAGE= 1
           VIS= all
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   HYDTrapezoid
           X1_Y1=           45.64   34.43  -146.50
           X128_Y1=         91.09   79.82   -91.36
           X1_Y128=        120.84    8.00   -91.36
           X128_Y128=       56.59    8.00  -146.50
           FIRSTSTAGE= 1
           SECONDSTAGE= 0
           THIRDSTAGE= 1
           VIS= all
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   HYDSquare
           X1_Y1=          49.1    66.08   -135.41
           X128_Y1=        -48.9   66.22   -135.41
           X1_Y128=        -48.8   135.51  -66.1
           X128_Y128=      49.2    135.36  -66.1
           FIRSTSTAGE= 1
           SECONDSTAGE= 0
           THIRDSTAGE= 1
           VIS= all
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   HYDDummyShape
           THETA= 90
           PHI= 90
           R= 100
           BETA= 0 0 0
           FIRSTSTAGE= 1
           SECONDSTAGE= 1
           THIRDSTAGE= 1
           VIS= all
\end{verbatim}

In order to declare a Hyde tracker detector in NPSimulation, the key
word {\it HydeTracker} should be specified in the geometry file. It
should then be followed by other keywords concerning the different
detectors present in the tracker. Such keywords available at the moment
are:

\begin{itemize}
   \item {HYDAnnular}
   \item {HYDTrapezoid}
   \item {HYDSquare}
   \item {HYDDummyShape}
\end{itemize}

Each keyword corresponds to a detector shape which have its own set of
keywords which is used to position the detector in the world volume
and to define the structure of the detector (basically the number of 
layers of the detector).

To position the detectors two possibilities exist. Either the cartesian 
coordinates (x,y,z) of each detector's corner are specified with the 
keywords {\it X1\_Y1}, {\it X1\_Y128}, {\it X128\_Y1} and {\it X128\_Y128}
(case of HYDTrapezoid), either the spherical coordinates of the detector's 
center are specified with the keywords {\it R}, {\it THETA} and {\it PHI} 
(case of HYDDummyShape). While the first solution is very helpful when 
working with the mechanical engineers, the second solution is useful when 
investigating new geometries.

In case of HYDAnnular only the position on the z-axis is given through
the keyword {\it Z}. Other keywords such as {\it RMIN} and {\it RMAX} 
are defined but have no effect. This should be changed in the future.

Concerning the structure of the detector, all the detectors have the 
possibility to have up to three layers of silicon. Each layer can be 
activated independently using the keywords {\it FIRSTSTAGE}, 
{\it SECONDSTAGE} and {\it THIRDSTAGE}. 


\subsection{Results of the simulation}
The results of the simulation are in the ROOT format and the output file 
is stored in the \$NPTool/Output/Simulation directory. The output ROOT file 
contains three classes:
\begin{itemize}
   \item {TInitialConditions:}
      This class records all the information concerning the event generator
      such as the vertex of interaction, the angles of emitted particles in 
      the center of mass and laboratory frames...
   \item {TInteractionCoordinates:}
      This class mainly records the cartesian and spherical coordinates of 
      interaction between a particle and a detector.
   \item {THydeTrackerData:}
      This class stores the results of the simulation for the tracker part 
      of the Hyde detector. Independently of the number and shape of the 
      detectors involved in the geometry, only {\it one} class is created for 
      the whole Hyde tracker detector. For each event the strips number 
      are recorded as well as the energy and time for the layers which are 
      involved in the telescope.
\end{itemize}


\subsection{Adding a new detector shape to Hyde}
A special class (HydeTrackerDummyShape) has been created to show how
to define a new module for the Hyde Tracker. This class describes a
simple 5x5 cm$^2$ square telescope made of three layers of silicon which  
has been used for some preliminary studies of the tracker. So, when 
considering adding a new module to the Hyde Tracker, please do not use
this class but create your own instead. However, for the explanations the 
HydeTrackerDummyShape case will be considered.

When adding a new detector you need to follow several steps:

\subsubsection{Definition of an index for the detector}
Since the results of the simulation are stored in an unique data class 
(HydeTrackerData) dealing with different module types, it is 
necessary to give an absolute number to each module. This is managed in
the HydeTrackerModule class where there is a map (m\_index) which
associates a name (the module type identifier) with an integer (the value 
of the index).

To add a new detector it is just needed to add in the {\it InitializeIndex()}
method a line such as:

\begin{verbatim}
   m_index["DummyShape"] = 1000;
\end{verbatim}

\subsubsection{Definition of the geometry and its readout}
This is done in the HydeTrackerDummyShape.\{h,cxx\} files. Concerning
the geometry it is defined in the {\it VolumeMaker()} method and the 
the positioning of the module is done in the {\it ConstructDetector()} 
method. 

Concerning the readout of the geometry it is done in the {\it ReadSensitive()} 
method and it is based on the method of the scorers available in G4. The 
scorers associated to the Hyde tracker are declared in the {\it 
InitializeScorers()} method.

\subsubsection{Definition of the scorers}
If the scorers are declared in the {\it InitializeScorers()} method they should
be defined in the HydeScorers.\{hh,cc\} files. All basic scorers to record
energy, time of flight and detector number are already implemented so when a new
detector is added to Hyde tracker there is nothing to add from this point
of vue. However scorers determining the strip number for the front and back
side of the silicon detector's first stage (if double-sided) should be implemented.
In case of the HydeTrackerDummyShape class this corresponds to the two classes
HYDScorerFirstStageFrontStripDummyShape and HYDScorerFirstStageBackStripDummyShape.

\subsubsection{Integration in HydeTracker}
In order to make the HydeTracker detector aware of the HydeTrackerDummyShape
module it has to be registered in the {\it ReadConfiguration()} method of the 
HydeTracker class. Don't forget to include the HydeTrackerDummyShape.hh 
header in the HydeTracker.cc file. Then, in the HydeTrackerDummyShape class 
the keywords used when the geometry file is read should be defined in the {\it 
ReadConfiguration()} method.


\section{NPAnalysis}
\subsection{General}
A set of general ROOT macros are available in the \$NPTool/NPAnalysis/macros 
directory. You can for example obtain some control plots about the shooting
conditions of the random variables. You can also calculate the geometrical 
efficiency of your setup.

The macros in this directory should be independant of the setup which is simulated.
Specific macros to Hyde tracker should be placed in the 
\$NPTool/NPAnalysis/Hyde/macros directory.


\subsection{Hyde}
The main analysis tool for the Hyde tracker is in the \$NPTool/NPAnalysis/Hyde
directory. For the moment the main feature is the reconstruction of the excitation 
energy.

\subsubsection{Running the analysis}
To run NPAnalysis the following command line should be executed:

\begin{verbatim}
   ./Analysis -E xxx.reaction -D yyy.detector RunToTreat.txt
\end{verbatim}

where xxx.reaction is the input file describing the event generator used in 
NPSimulation and yyy.detector is the input file describing the detector geometry
used in NPSimulation. All these input files are based on keywords and can be found 
in the \$NPTool/Inputs subdirectories. The RunToTreat.txt file contains the
name of the files (either from NPSimulation or from real experiment) which should
be analysed. The name of the tree should also be specified. An example 
of such a file is given here:

\begin{verbatim}
   TTreeName
           SimulatedTree
   RootFileName
   %       ../../Outputs/Simulation/mySimul.root
           ../../Outputs/Simulation/132Sndp_3p3_10MeVA_T0_B1_E0_S05mm.root
   %       ../../Outputs/Simulation/134Snpt_1h9_10MeVA_T1_B1_E0_S05mm.root
\end{verbatim}


\subsubsection{Results of the analysis}
The results of the anaysis are stored in a ROOT file in the \$NPTool/Output/Analysis
directory. For the moment the main feature available is the reconstructionof the 
excitation energy.


\subsubsection{Structure of the analysis}
******* to be documented *********


\end{document}

